---
title: "scRNAseq comparison of different models of ZFTA::RELA tumors"
author: "Sara Danielli"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
    toc_float: yes
---

```{r, setup, include = FALSE}
library(knitr)
opts_chunk$set(
  echo = TRUE, cache = TRUE, warning = FALSE, comment = FALSE)
```

```{r message=FALSE, warning=FALSE}
# Load packages -----------------------------------
library(dplyr)
library(Seurat)
library(ggplot2)
library(patchwork)
library(SeuratObject)
library(data.table)
library(tidyverse)
library(future)
#library(clustree)
#library(clusterProfiler)
library(ggpubr)
library(cowplot)
library(ggrastr)
library(readxl)
library(writexl)

# Organize environment  -----------------------------------
base_dir <- "/Users/sdaniell/Dropbox (Partners HealthCare)/Sara Danielli/Project/Ependymoma/6 - scRNAseq models"
analysis_dir <- file.path(base_dir, 'analysis')
resource_dir <- file.path(base_dir, 'scripts/resources')

plot_dir <- file.path(analysis_dir, 'plots')
if (!dir.exists(plot_dir)){dir.create(plot_dir, recursive = T)}

data_dir <- file.path(analysis_dir, 'data')

source(file.path(resource_dir, "single_cell_preprocessing_helper_functions.R"))
source(file.path(resource_dir, 'Plot_style.R'))


# Define color palettes  -----------------------------------
col_model = c('grey80', paletteer::paletteer_d("beyonce::X18")[3:5])
colors_metaprograms<- c("gray30","#F99E93FF","#9E5E9BFF","#74ADD1FF",'#0F4F8B', "#ACD39EFF","#96410EFF", 'mistyrose1')

names(colors_metaprograms) <- c("Cycling", "Neuroepithelial-like", "Radial-glia-like", 
                          "Embryonic-neuronal-like", "Neuronal-like" ,"Ependymal-like", "MES-like", "Embryonic-like")


# load Seurat object -----------------------------------
seurat_obj_merged <- readRDS(file.path(analysis_dir, "data/Patient_models_merged.rds"))


```

# UMAP plots

Plot UMAP colored by different metadata columns

## By cluster
```{r UMAP clusters, fig.width = 5, fig.height = 4, fig.align="center"}
## by clusters
DimPlot(seurat_obj_merged, group.by = 'seurat_clusters', reduction = "umap", pt.size = 1)
ggsave(file.path(plot_dir, "2_UMAP_clusters.pdf"), width=6, height=5)
```
## By model
```{r UMAP model, fig.width = 5, fig.height = 4, fig.align="center"}
## by model
DimPlot(seurat_obj_merged, reduction = "umap", group.by = 'model', pt.size = 1, cols = col_model)
ggsave(file.path(plot_dir, "2_UMAP_model.pdf"), width=6, height=5)

```  

## By metaprogram
```{r UMAP metaprogram, fig.width = 12, fig.height = 4, fig.align="center"}
## by metaprogram
p1 <- DimPlot(seurat_obj_merged, reduction = "umap", group.by = 'SingleR.labels', pt.size = 1, cols = colors_metaprograms)
p2 <-  DimPlot(seurat_obj_merged, reduction = "umap", group.by = 'signature_1', pt.size = 1, cols = colors_metaprograms)
(p1 + p2)
ggsave(file.path(plot_dir, "2_UMAP_metaprogram.pdf"), width=12, height=5)
```  

## By scores
```{r UMAP scores, fig.width = 9, fig.height = 5, fig.align="center"}
scores <- c('Neuronal-like', 'MES-like', 'Radial-glia-like', 'Ependymal-like', 'Neuroepithelial-like', 'Cycling')
names(scores) <- c('Neuronal-like', 'MES-like', 'Radial-glia-like', 'Ependymal-like', 'Neuroepithelial-like', 'Cycling')
p <- FeaturePlot(seurat_obj_merged, features = scores, reduction = "umap", order = TRUE, combine = FALSE, pt.size = 1)

for(i in 1:length(p)) {
  p[[i]] <- p[[i]] + NoAxes() + paletteer::scale_colour_paletteer_c("viridis::mako", direction = -1) +
      labs (title = names(scores)[[i]]) 
  }

cowplot::plot_grid(plotlist = p, ncol=3)
ggsave(file.path(plot_dir, "3_UMAP_scores.pdf"), width=9, height=5, dpi=300)
```   

# Violin plots
Plot Violin plots of tumor program scores 

```{r Vln plots, fig.width = 8, fig.height = 8, fig.align="center"}
p <- VlnPlot(seurat_obj_merged, features = scores, group.by = 'model', combine = FALSE, pt.size=0, cols = col_model) 

for(i in 1:length(p)) {
    p[[i]] <- p[[i]] + NoLegend() +
      labs (y='Module score, AU', x='') + 
      scale_fill_manual(values = col_model) + 
      geom_boxplot(outlier.shape=NA, width=0.1, fill="white") +
      # stat_compare_means(aes(label = after_stat(p.signif)), 
      #                    method = 't.test', 
      #                    size = 5, 
      #                    label.y.npc = 0.92, 
      #                    label.x.npc = 0.5, 
      #                    ) + 
      NoLegend()
      theme_vln
  }
plot_grid(plotlist = p, ncol=3)
ggsave(file.path(plot_dir, "4_VlnPlot_scores.pdf"), width=6, height=8)
```

Plot Violin plots of selected scores 

```{r Vln plots 2, fig.width = 8, fig.height = 4, fig.align="center"}
plot_grid(p[[5]], p[[1]], p[[4]], p[[3]], ncol = 4) 
ggsave(file.path(plot_dir, "5_VlnPlot_scores_selected.pdf"), width=8, height=4)
```

# Violin plots
Plot Violin plots of markers 

```{r Vln plot markers, fig.width = 8, fig.height = 6, fig.align="center"}
markers <- c('VIM', 'NES', 'GRID2', 'DCX', 'TUBB3', 'CCDC40', 'S100B')

p <- VlnPlot(seurat_obj_merged, features = markers, group.by = 'model', combine = FALSE, pt.size=0, cols = col_model) 

for(i in 1:length(p)) {
    p[[i]] <- p[[i]] + NoLegend() +
      labs (y='Module score, AU', x='') + 
      scale_fill_manual(values = col_model) + 
      geom_boxplot(outlier.shape=NA, width=0.1, fill="white") +
      # stat_compare_means(aes(label = after_stat(p.signif)), 
      #                    method = 't.test', 
      #                    label.x.npc = 0.5, 
      #                    label.y.npc = 0.92, 
      #                    label.x = 1.3
      #                    ) + 
            theme_vln +
      NoLegend()
  }
plot_grid(plotlist = p, ncol=4)
ggsave(file.path(plot_dir, "5_VlnPlot_markers.pdf"), width=8, height=6)
```


# Barplot 

Barplot of label transfer
```{r barplot singleR, fig.width = 10, fig.height = 4, fig.align="center"}
## Bar plot
p1 <- plot_bar(seurat_obj_merged, seurat_obj_merged$model, seurat_obj_merged$SingleR.labels, col = colors_metaprograms) 
p2 <- plot_bar(seurat_obj_merged, seurat_obj_merged$model, seurat_obj_merged$signature_1, col = colors_metaprograms) 
(p1 + p2)
ggsave(file.path(plot_dir, "7_Barplot_label_transfer_model.pdf"), width=10, height=4)
```


Number of cells assigned to cell states
```{r table nr cells NMF scoring, fig.width = 4, fig.height = 4, fig.align="center"}
# Visualize the number of cells  per sample considering quality
metadata_seurat <- seurat_obj_merged@meta.data
# Print out number of good and bad quality cells
number_cells <- metadata_seurat %>% 
  group_by(model, signature_1) %>% 
  summarise(n = n())  %>%
  pivot_wider(names_from = signature_1, values_from = n) 

```

Barplot of label transfer
```{r barplot singleR split by patient, fig.width = 10, fig.height = 4, fig.align="center"}
## Bar plot
plot_bar(seurat_obj_merged, seurat_obj_merged$FileName, seurat_obj_merged$signature_1, col = colors_metaprograms) 
ggsave(file.path(plot_dir, "7_Barplot_label_transfer_sample.pdf"), width=8, height=4)
```

# Cell state plot

Plot distribution of samples on the hierarchy plot
```{r cellstate plot 4quadrant colored by metaprogram, fig.width = 5.5, fig.height = 4, fig.align="center"}
x1 = "Neuroepithelial-like"
x2 = "Embryonic-like"
y1 = "Neuronal-like"
y2 = "Ependymal-like"

group.by="model"
sample=seurat_obj_merged
variables_to_retrieve <- c(x1, x2, y1,y2)

# And store them as a tibble.
scores <- sample@meta.data[, variables_to_retrieve]
# Shuffle the cells so that we accomplish a random plotting, not sample by sample.

# Compute Y axis values.
d <- apply(scores, 1, function(x){max(x[c(x1, x2)]) - max(x[c(y1, y2)])})

# Compute X axis values.
x <- vapply(seq_along(d), function(x) {
  if (d[x] > 0) {
    d <- log2(abs(scores[x, x1] - scores[x, x2]) + 1)
    ifelse(scores[x, x1] < scores[x, x2], d, -d)
  } else {
    d <- log2(abs(scores[x, y1] - scores[x, y2]) + 1)
    ifelse(scores[x, y1] < scores[x, y2], d, -d)
  }
}, FUN.VALUE = numeric(1))

names(x) <- rownames(scores)

# Plot.
df_E <- data.frame(row.names = rownames(scores))
df_E[["set_x"]] <- x
df_E[["set_y"]] <- d
df_E[["group.by"]] <- sample@meta.data[, group.by]
df_E[["subtype"]] <- seurat_obj_merged$model

col_model
names(col_model) <- c('Patient', 'PDX', 'adherent', 'spheres')

ggplot(df_E,aes(x=set_x,
              y=set_y,
              col=group.by))+
  geom_point(size=1)+
  scale_color_manual(values=col_model)+
   theme_cellstate_plot + 
  labs(x = 'Relative meta-module score', y = 'Relative meta-module score', title = 'ST-EPN cellular hierarchy', subtitle = 'n = 3,941 cells') +
    geom_vline(xintercept = 0, linetype="dotted") + geom_hline(yintercept = 0, linetype="dotted") 

ggsave(file.path(plot_dir, "8_CellState_plot_4way_model.pdf"), width=5.5, height=4)

```
